Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2016-01-31 ijar-npb-ms.tex
Edits to Section 7 and 8 -- feel free to roll back any/all of them!

\section{Computations \& Examples}

@@

The methods of this paper have been implemented in the \textbf{R}
\citep{R} package \texttt{ReliabilityTheory} \citep{2015:aslett-RT},
providing an easy to use interface for reliability practitioners.  The
primary function, which computes the upper and lower posterior
predictive system survival probabilities as in \eqref{eq:LwrSysPost}, is
named \texttt{nonParBayesSystemInferencePriorSets()}.  The user
specifies the times at which to evaluate the bounds, the survival
signature ($\Phi(\cdot)$), the component test data ($\vec{t}^1, \ldots,
\vec{t}^K$), and the prior parameter set for each component type and
time ($\PtZi{k}$, via $\nzu_{k,t}, \nzl_{k,t}, \yzu_{k,t}$, and
$\yzl_{k,t}$).  All computations of $\lRsys$
and $\uRsys$ at different time points are performed in parallel
automatically where the CPU has multiple cores.

Note that computation of the system signature itself can be simplified by
expressing the structure of the system as an undirected graph using the
\texttt{computeSystemSurvivalSignature()} function in the same package,
leaving only data and prior to be handled.  These publicly available
functions have been used in computing all the following examples for
reproducibility.  \ref{ap:software} details how to use this software.

@@

@@

In Figure~\ref{fig:bridge-fitting}, 
test data for component type T3 is $\vec{t}^3 = (0.5, 1.5, 2.5, 3.5)$,
and so in line with expectations.
The posterior set of reliability functions for each component type and the whole system
is considerably smaller as compared to the prior set
(due to the low prior strength intervals
$[\ntzl{1},\ntzu{1}] = [\ntzl{2},\ntzu{2}] = [1,2]$, $[\ntzl{3},\ntzu{3}] = [1,4]$)
and so giving more precise reliability statements.
We see that posterior lower and upper functioning probabilities drop at those times $t$
when there is a failure time in the test data,
or a drop in the prior functioning probability bounds.
Note that the lower bound for prior system reliability function is zero
due to the prior lower bound of zero for T1;
for the system to function, at least two components of type T1 must function.

@@

In Figure~\ref{fig:bridge-early},
test data of component type T3 is instead $\vec{t}^3 = (0.6, 0.7, 0.8, 0.9)$,
and so earlier than expected.
Compared to Figure~\ref{fig:bridge-fitting},
posterior functioning intervals for T3 are wider between $t=1$ and $t=3.5$,
reflecting additional imprecision due to prior-data conflict.
For $t > 1$, it is clearly visible how $\ytnu{3}$ is halfway between $\ytzu{3}$ and $s_t^3/n_3 = 0$
(weights $\ntzu{3} = 4$ and $n_3 = 4$),
while $\ytnl{3}$ is one-fifth of $\ytzl{3}$
(weights $\ntzl{3} = 1$ and $n_3 = 4$).
Note that the posterior system functioning probability is constant for $t \in [1,2]$
because the prior functioning probability is constant for $t \in [1,2]$
and there are no failure times $\in [1,2]$ in the test data.

@@

In Figure~\ref{fig:bridge-late}, 
test data of component type T3 is now $\vec{t}^3 = (4.1, 4.2, 4.3, 4.4)$,
and so observed failures are later than expected.
Here we see that for $t \in [2,4]$,
posterior functioning bounds for T3 are even wider than prior functioning bounds.
The width turns back to being half the prior width
only after the four failures.
The imprecision carries over to the system bounds,
where we see wider bounds as compared to the other two scenarios
especially between $t=2$ and $t=4$.

@@

As example to illustrate the potential of our model,
we consider a simplified automotive brake system.
The master brake cylinder (M) activates all four wheel brake cylinders (C1 -- C4),
which in turn actuate a braking pad assembly each (P1 -- P4).
The hand brake mechanism (H) goes directly to the brake pad assemblies P3 and P4;
the car brakes when at least one brake pad assembly is actuated.
All values for $\Phi \not\in \{0,1\}$ are given in Table~\ref{tab:brake-survsign}.
\begin{table}
\centering
\begin{tabular}{cccclcccccl}
  \toprule
M & H & C & P & $\Phi$ & \quad & M & H & C & P & $\Phi$\\ 
  \midrule
1 & 0 & 1 & 1 & 0.25 & & 1 & 0 & 2 & 1 & 0.50 \\ 
1 & 0 & 1 & 2 & 0.50 & & 1 & 0 & 2 & 2 & 0.83 \\ 
1 & 0 & 1 & 3 & 0.75 & & 1 & 0 & 3 & 1 & 0.75 \\ 
0 & 1 & 0 & 1 & 0.50 & & 1 & 1 & 0 & 1 & 0.50 \\ 
0 & 1 & 0 & 2 & 0.83 & & 1 & 1 & 0 & 2 & 0.83 \\ 
0 & 1 & 1 & 1 & 0.62 & & 1 & 1 & 1 & 1 & 0.62 \\ 
0 & 1 & 1 & 2 & 0.92 & & 1 & 1 & 1 & 2 & 0.92 \\ 
0 & 1 & 2 & 1 & 0.75 & & 1 & 1 & 2 & 1 & 0.75 \\ 
0 & 1 & 2 & 2 & 0.97 & & 1 & 1 & 2 & 2 & 0.97 \\ 
0 & 1 & 3 & 1 & 0.88 & & 1 & 1 & 3 & 1 & 0.88 \\ 
   \bottomrule
\end{tabular}
\caption{Survival signature values $\not\in \{0,1\}$ for the simplified automotive brake system depicted in Figure~\ref{fig:brakesystem}.
}
\label{tab:brake-survsign}
\end{table}
The system layout is depicted in Figure~\ref{fig:brakesystem},
together with prior and posterior sets of reliability functions for the four component types and the complete system.
Observed lifetimes from test data are indicated by tick marks in each of the four component type panels,
where $n_\text{M}=5$, $n_\text{H}=10$, $n_\text{C}=15$, and $n_\text{P}=20$.
We assume $[\ntzl{\text{M}},\ntzu{\text{M}}] = [1,8]\ \forall t$,
and $[\nktzl, \nktzu] = [1,2]$ for $k \in \{\text{H, C, P}\}$ and all $t$.
Prior functioning probability bounds for M are based on
a Weibull cdf with shape $2.5$ and scales $6$ and $8$ for the lower and upper bound, respectively.
The prior bounds for P can be seen as the least committal bounds
derived from an expert statement of $\ytz{\text{P}} \in [0.5, 0.65]$ for $t=5$ only.

***We see that posterior lower and upper functioning probabilities drop at those times $t$
for which there is a failure time in the test data,
or a drop in the prior functioning probability bounds.
The lower bound for prior system reliability drops to zero at $t=5$
since the prior lower bound for P drops to zero at $t=5$;
for the system to function, at least one brake pad assembly must function.***

@@

@@

***Some extra illustrations? E.g. the effect of extra redundancy on $\Rsys(t)$ intervals like in Risk Analysis paper?
(could nicely illustrate challenges in decision making with intervals:
do I prefer adding redundancy such that there is a chance for a large effect, but high uncertainty for it,
or do I rather add redundancy such that effect is smaller but more certain?***

***Other interesting extra illustration:
show effect of replacement of components in system at certain time points
(could be corrective (component failed) or preventive replacement),
replaced components constitute a new type with shifted component reliability function***

\section{Conclusions and Outlook}

Upscaling the survival signature to large real-world systems and networks, consisting of thousands of components, is a major challenge.
However, even for such systems the fact that one only needs to derive the survival signature once for a system is an advantage,
and also the monotonicity of the survival signature for coherent systems is very useful if one can only derive it partially.

The survival signature and its use for uncertainty quantification for system
reliability can be generalized quite straightforwardly, mainly due to the simplicity of this concept.
For example, one may generalize the system structure function from a binary function to a probability, %\citep[see]{CCM16},
to reflect uncertainty about system functioning for known states of its components, with a further
generalization to imprecise probabilities possible. It is also conceptually easy to generalize the
survival signature to multi-state systems such that it again summarizes the structure function
in a manner that is sufficient for a range of uncertainty quantifications for the system reliability.

***Extend the model to deal with right-censored observations which are common in the reliability setting.
The minimal assumption that a component can either fail immediately after censoring or live forever
will be simple to implement but will lead to high imprecision,
whereas assuming exchangeability with other surviving components at moment of censoring
will be more complex to accomodate but will lead to less imprecision.
This excangeability assumption lies at the core of the Kaplan-Meier estimator \citep{1958:kaplan-meier},
and has already been adopted by \citet{2004:coolen-yan} in an imprecise probability context.

***right-censoring will also allow to use data from running system,
to calculate its remaining useful life (RUL).

>>>>

\section{Practical Usage and Examples}

@@

The methods of this paper have been implemented in the \textbf{R}
\citep{R} package \texttt{ReliabilityTheory} \citep{2015:aslett-RT},
providing an easy to use interface for reliability practitioners.  The
primary function, which computes the upper and lower posterior
predictive system survival probabilities as in \eqref{eq:LwrSysPost}, is
named \texttt{nonParBayesSystemInferencePriorSets()}.  The user
specifies the times at which to evaluate the bounds, the survival
signature ($\Phi(\cdot)$), the component test data ($\vec{t}^1, \ldots,
\vec{t}^K$), and the prior parameter set for each component type and
time ($\PtZi{k}$, via $\nzu_{k,t}, \nzl_{k,t}, \yzu_{k,t}$, and
$\yzl_{k,t}$).  All computations of $\lRsys$
and $\uRsys$ at different time points are performed in parallel
automatically where the CPU has multiple cores and making automatic
use of the theoretical results in Section \ref{sec:setsofrel} where
applicable, performing exhaustive search in the few cases they are not.

Note that computation of the system signature itself can be simplified by
expressing the structure of the system as an undirected graph using the
\texttt{computeSystemSurvivalSignature()} function in the same package,
leaving only data and prior to be handled.  These publicly available
functions have been used in computing all the following examples for
reproducibility.  See Appendix \ref{ap:software} for further details of how to use this software.

@@

\subsubsection{Toy example}

@@

In Figure~\ref{fig:bridge-fitting}, 
test data for component type T3 is $\vec{t}^3 = (0.5, 1.5, 2.5, 3.5)$,
and so in line with expectations.
The posterior set of reliability functions for each component type and the whole system
is considerably smaller compared to the prior set
(due to the low prior strength intervals
$[\ntzl{1},\ntzu{1}] = [\ntzl{2},\ntzu{2}] = [1,2]$, $[\ntzl{3},\ntzu{3}] = [1,4]$)
and so giving more precise reliability statements.
We see that posterior lower and upper functioning probabilities drop at those times $t$
when there is a failure time in the test data,
or a drop in the prior functioning probability bounds.
Note that the lower bound for the prior system reliability function is zero
due to the prior lower bound of zero for T1;
for the system to function, at least two components of type T1 must function.

@@

In Figure~\ref{fig:bridge-early},
test data of component type T3 is instead $\vec{t}^3 = (0.6, 0.7, 0.8, 0.9)$,
and so earlier than expected.
Compared to Figure~\ref{fig:bridge-fitting},
posterior functioning intervals for T3 are wider between $t=1$ and $t=3.5$,
reflecting additional imprecision due to prior-data conflict.
For $t > 1$, it is clearly visible how $\ytnu{3}$ is halfway between $\ytzu{3}$ and $s_t^3/n_3 = 0$
(weights $\ntzu{3} = 4$ and $n_3 = 4$),
while $\ytnl{3}$ is one-fifth of $\ytzl{3}$
(weights $\ntzl{3} = 1$ and $n_3 = 4$).
Note that the posterior system functioning probability is constant for $t \in [1,2]$
because in that interval the prior functioning probability is constant
and there are no observed failures.

@@

In Figure~\ref{fig:bridge-late}, 
test data of component type T3 is now $\vec{t}^3 = (4.1, 4.2, 4.3, 4.4)$,
and so observed failures are later than expected.
Here we see that for $t \in [2,4]$,
posterior functioning bounds for T3 are even wider than prior functioning bounds.
The width turns back to being half the prior width
only after the four failures.
The imprecision carries over to the system bounds,
where we see wider bounds as compared to the other two scenarios
especially between $t=2$ and $t=4$.  In particular, also note
that at the system level prior data conflict ceases after
$t \approx 2.6$, although the prior data conflict for the 
individual component type T3 extends well beyond $t=4$.  This 
demonstrates the power of this technique to identify prior
data conflict which is actually consequential at the system 
level, not just the component level --- in other words, for 
mission times $t>2.6$, we can diagnose that the prior data conflict 
need not be of concern for this system viewed as a whole.

@@

\subsubsection{Automotive brake system}

We also consider a simplified automotive brake system.
The master brake cylinder (M) activates all four wheel brake cylinders (C1 -- C4),
which in turn actuate a braking pad assembly each (P1 -- P4).
The hand brake mechanism (H) goes directly to the brake pad assemblies P3 and P4;
the car brakes when at least one brake pad assembly is actuated.
All values for $\Phi \not\in \{0,1\}$ are given in Table~\ref{tab:brake-survsign}.
\begin{table}
\centering
\begin{tabular}{cccclcccccl}
  \toprule
M & H & C & P & $\Phi$ & \quad & M & H & C & P & $\Phi$\\ 
  \midrule
1 & 0 & 1 & 1 & 0.25 & & 1 & 0 & 2 & 1 & 0.50 \\ 
1 & 0 & 1 & 2 & 0.50 & & 1 & 0 & 2 & 2 & 0.83 \\ 
1 & 0 & 1 & 3 & 0.75 & & 1 & 0 & 3 & 1 & 0.75 \\ 
0 & 1 & 0 & 1 & 0.50 & & 1 & 1 & 0 & 1 & 0.50 \\ 
0 & 1 & 0 & 2 & 0.83 & & 1 & 1 & 0 & 2 & 0.83 \\ 
0 & 1 & 1 & 1 & 0.62 & & 1 & 1 & 1 & 1 & 0.62 \\ 
0 & 1 & 1 & 2 & 0.92 & & 1 & 1 & 1 & 2 & 0.92 \\ 
0 & 1 & 2 & 1 & 0.75 & & 1 & 1 & 2 & 1 & 0.75 \\ 
0 & 1 & 2 & 2 & 0.97 & & 1 & 1 & 2 & 2 & 0.97 \\ 
0 & 1 & 3 & 1 & 0.88 & & 1 & 1 & 3 & 1 & 0.88 \\ 
   \bottomrule
\end{tabular}
\caption{Survival signature values $\not\in \{0,1\}$ for the simplified automotive brake system depicted in Figure~\ref{fig:brakesystem}.
}
\label{tab:brake-survsign}
\end{table}
The system layout is depicted in Figure~\ref{fig:brakesystem},
together with prior and posterior sets of reliability functions for the four component types and the complete system.
Observed lifetimes from test data are indicated by tick marks in each of the four component type panels,
where $n_\text{M}=5$, $n_\text{H}=10$, $n_\text{C}=15$, and $n_\text{P}=20$.
We assume $[\ntzl{\text{M}},\ntzu{\text{M}}] = [1,8]\ \forall t$,
and $[\nktzl, \nktzu] = [1,2]$ for $k \in \{\text{H, C, P}\}$ and all $t$.
Prior functioning probability bounds for M are based on
a Weibull cdf with shape $2.5$ and scales $6$ and $8$ for the lower and upper bound, respectively.
The prior bounds for P can be seen as the least committal bounds
derived from an expert statement of $\ytz{\text{P}} \in [0.5, 0.65]$ for $t=5$ only.

@@

It is also interesting to note that the prior data conflict 
which is consequential at the system level occurs over 
roughly the same range in $t$ as there is prior data 
conflict for component type P.  Indeed, this occurs despite there being 
prior data conflict in both M and C over much larger ranges, 
giving valuable insight into which prior requires further 
expert attention --- thus the technique avoids wasted time 
addressing prior data conflict in components which may not 
be relevant when propagated to the uncertainty in the whole system.

@@

%***Some extra illustrations? E.g. the effect of extra redundancy on $\Rsys(t)$ intervals like in Risk Analysis paper?
%(could nicely illustrate challenges in decision making with intervals:
%do I prefer adding redundancy such that there is a chance for a large effect, but high uncertainty for it,
%or do I rather add redundancy such that effect is smaller but more certain?***
%
%***Other interesting extra illustration:
%show effect of replacement of components in system at certain time points
%(could be corrective (component failed) or preventive replacement),
%replaced components constitute a new type with shifted component reliability function***

\section{Conclusions}

In this paper we have contributed an imprecise Bayesian nonparametric approach
to system reliability with multiple types of components.
The approach allows modeling partial or imperfect prior knowledge on component failure distributions
in a flexible way through bounds on the functioning probability for a grid of time points, 
and combines this information with test data in an imprecise Bayesian framework.
Component-wise predictions on the number of functioning components
are then combined to bounds for the system survival probability by means of the survival signature.
New results on first-order stochastic dominance for the Beta-Binomial distribution
enable closed-form solutions for these bounds in most cases and 
avoid exponential growth in the complexity of computing the estimate
as the number of components grows.
The widths of the resulting system reliability bounds
reflect the amount of test data, the precision of prior knowledge,
and crucially provide an easily used method to identify 
whether these two information sources are in conflict in a
way which is of consequence to the whole system reliability
estimate.

These methodological contributions can be immediately used in
applications by reliability practitioners as a result of the 
easy to use software tools which have been provided.

%% LOUIS: do we really want outlook?  I am inclined not to
%% discuss this in a non-review/tutorial paper because we
%% don't want to alert others and get scooped on future
%% ideas!
%Upscaling the survival signature to large real-world systems and networks, consisting of thousands of components, is a major challenge.
%However, even for such systems the fact that one only needs to derive the survival signature once for a system is an advantage,
%and also the monotonicity of the survival signature for coherent systems is very useful if one can only derive it partially.  
%
%The survival signature and its use for uncertainty quantification for system
%reliability can be generalized quite straightforwardly, mainly due to the simplicity of this concept.
%For example, one may generalize the system structure function from a binary function to a probability, %\citep[see]{CCM16},
%to reflect uncertainty about system functioning for known states of its components, with a further
%generalization to imprecise probabilities possible. It is also conceptually easy to generalize the
%survival signature to multi-state systems such that it again summarizes the structure function
%in a manner that is sufficient for a range of uncertainty quantifications for the system reliability. 
%
%
%***Extend the model to deal with right-censored observations which are common in the reliability setting.
%The minimal assumption that a component can either fail immediately after censoring or live forever
%will be simple to implement but will lead to high imprecision,
%whereas assuming exchangeability with other surviving components at moment of censoring
%will be more complex to accomodate but will lead to less imprecision.
%This excangeability assumption lies at the core of the Kaplan-Meier estimator \citep{1958:kaplan-meier},
%and has already been adopted by \citet{2004:coolen-yan} in an imprecise probability context.
%
%***right-censoring will also allow to use data from running system,
%to calculate its remaining useful life (RUL).
%
